---
title: "EB-rMAP Prior with a Time-to-Event Endpoint"
author: "Hongtao Zhang"
date: "12/28/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EB-rMAP Prior with a Time-to-Event Endpoint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we showcase how the EB-rMAP method can be implemented when a time-to-event (TTE) endpoint is of interest. It is essentially replicating the data analysis carried out in section 4 of Zhang et al. [1]. 

## Data
The data in 10 oncology studies come from Roychoudhury and Neueuschwander (2020) [2]. In each study, the four-year follow-up period was partitioned into 12 intervals, and the number of events and total exposure (follow-up) time in years were recorded. For illustration purposes, the first 9 trials are regarded as historical trials, while the 10th trial is considered the current trial. The full data is displayed in the Appendix. We aggregate the data between year 0 and 1.5 to implement the EB-rMAP method for the log hazard rates. 
```{r, warning=F}
library(RBesT)
library(tidyverse)
library(knitr)
library(kableExtra)
library(EBrmap)

set.seed(712)

Time <- c("0.00-0.25", "0.25-0.50", "0.50-0.75", "0.75-1.00",
          "1.00-1.25", "1.25-1.50", "1.50-1.75", "1.75-2.08",
          "2.08-2.50", "2.50-2.92", "2.92-3.33", "3.33-4.00")
FIOCCO.n.events <- c(1,  3,  3,  4,  3,  0,  0,  2,  0,  6,  0,  0,  9,  1,  0, 10,  6,  6,  5,  9,
                     9,  3,  0,  0,  1,  3,  5,  7,  9,  4,  5, 10,  0,  0,  3,  7,  1,  2,  2,  4, 
                     3,  1,  3,  0,  0,  0,  0,  0,  5,  3,  6,  2,  3,  3,  0,  2,  1,  1,  1,  0, 
                     0,  6,  3, 12,  8,  2, 3,  2, 11,  1,  0, 10,  2,  2,  5,  3,  3,  3,  2,  3, 
                     3,  0,  0,  0,  0,  1,  3,  4, 1,  1,  4,  1,  6,  0,  0,  0,  2,  0,  3,  1, 
                     4,  0,  1,  1,  0,  0,  0,  0,  1,  5, 17, 0,  2,  7,  8,  4,  0,  6,  2,  0)
FIOCCO.exp.time <- c(9.4,  8.8,  7.9,  7.0,  6.1,  5.8,  5.8,  7.3,  8.8,  7.6,  6.2, 10.0, 21.1, 
                     19.9, 19.8, 18.5, 16.5, 15.0, 13.6, 15.7, 16.2, 13.6, 12.5, 20.1, 21.9, 21.4, 
                     20.4, 18.9, 16.9, 15.2, 14.1, 16.2, 18.5, 18.3, 17.0, 24.5,  5.6,  5.2,  4.8, 
                     4.0,  3.1,  2.6,  2.1,  2.3,  2.9,   2.9,  2.9,  4.7,  6.4,  5.4,  4.2,  3.2, 
                     2.6,  1.9,  1.5,  1.7, 1.5,  1.0,  0.6,  0.7, 17.8, 17.0, 15.9, 14.0, 11.5, 
                     10.2,  9.6, 11.9, 12.4,  9.9,  9.4, 12.1,  8.0,  7.5,  6.6,  5.6,  4.9,  4.1, 
                     3.5,  3.8,  3.6,  2.9,  2.9,  4.7,  9.2,  9.1,  8.6,  7.8,  7.1,  6.9,  6.2,  
                     7.4,  8.0,  6.7,  6.6, 10.7,  5.2,  5.0,  4.6,  4.1,  3.5,  3.0,  2.9,  3.5, 
                     4.2,  4.2,  4.1,  6.7, 23.4, 22.6, 19.9, 17.8, 17.5, 16.4, 14.5, 17.2, 21.0, 
                     19.7, 17.4, 27.5)


rmat <- 
  as_tibble(matrix(FIOCCO.n.events, nrow = 12, ncol = 10, byrow = F), 
            .name_repair = "minimal")
colnames(rmat) <- c(paste("Hist",1:9, sep=""), "Curr")
rdt <- tibble(Interval=Time, rmat)

Emat <- 
  as_tibble(matrix(FIOCCO.exp.time, nrow = 12, ncol = 10, byrow = F), 
            .name_repair = "minimal")
colnames(Emat) <- c(paste("Hist",1:9, sep=""), "Curr")
Edt <- tibble(Interval=Time, Emat)

rvec <- rdt[1:6,] %>% select(-c("Interval")) %>% colSums()
Evec <- Edt[1:6,] %>% select(-c("Interval")) %>% colSums()
adt <- tibble(Trial=c(paste("Hist",1:9, sep=""), "Curr"), NEvent=rvec, TotExp=Evec)

adt %>% kbl(caption="Aggregated Data Through Year 1.5") %>% 
  kable_classic(full_width = F, html_font = "Cambria", latex_options = "HOLD_position")
```

The random effect meta-analysis yields a point estimate of 0.38, and the 95% CI is (0.282, 0.510). 
```{r, eval=FALSE}
meta::metarate(NEvent, TotExp, Trial, data=adt[1:9,])
```

## MAP Prior and EB-rMAP Prior
The MAP prior is derived using the *`RBesT::gMAP`* command via a Poisson regression. It is critical to specify the log total exposure as the offset of the model. We use the same specifications in section 4 of Zhang et al. [1]: a $N(0,10^2)$ hyper-prior for the overall mean and a $HN(0.5)$ hyper-prior for the exchangeability parameter $\tau$. The MAP prior is then approximated by a mixture of two Gamma distributions (mean and number of observations, or `mn` parameterization). The vague prior to construct the robust MAP prior is $Gamma(0.38,1)$, which has the same mean with the meta-analysis but the effective sample size is only 1. 
```{r}
histdt <- adt[1:9,]

map_mcmc <- gMAP(NEvent ~ 1 + offset(log(TotExp)) | Trial, data = histdt, family = poisson, 
                 tau.dist = "HalfNormal", tau.prior = cbind(0, 0.5),
                 beta.prior=cbind(0, 10))
(map_hat <- automixfit(map_mcmc))

(vague_prior <- mixgamma(c(1, 0.38, 1), param = "mn"))
```

Suppose that we are designing a new single arm trial, augmented by the historical data via the robust MAP prior. The projected total exposure in the current trial is 117.6 years. Note that there are various software packages that may facilitate estimating this quantity via simulation, such as EAST and [*simtrial*](https://merck.github.io/simtrial/index.html) package. We examine the behavior of EB-rMAP weight correspoding to $\gamma =$ 0.85, 0.9 and 0.95 respectively. 
```{r}
y_range <- c(0, 50)
n <- 117.6

obj1 <- EB_rMAP(map_hat, vague_prior, 0.85, n, y_range)
obj2 <- EB_rMAP(map_hat, vague_prior, 0.9, n, y_range)
obj3 <- EB_rMAP(map_hat, vague_prior, 0.95, n, y_range)

plotdt1 <- obj1$wdt %>% mutate(Gamma=obj1$ppp_cut)
plotdt2 <- obj2$wdt %>% mutate(Gamma=obj2$ppp_cut)
plotdt3 <- obj3$wdt %>% mutate(Gamma=obj3$ppp_cut)
plotdt <- rbind(plotdt1, plotdt2, plotdt3)
```
```{r, fig.width=6, fig.fullwidth=TRUE}
plotdt %>% 
  ggplot(aes(x=y, y = w_eb, color=factor(Gamma))) + geom_line(size=1) +
  xlab("Observed Number of Events") + ylab("EB-rMAP Weight") + theme_bw() +
  geom_vline(xintercept=0.38*n, linetype="dashed") +
  scale_color_discrete(name="Gamma")
```

We can immediately see that the choice of $\gamma=0.95$ might be too stringent, as the weight is close to 1 even when the observed hazard rate in the current trial is close to historical mean hazard rate. After the initial screen, both $\gamma=0.9$ and 0.85 appear to be reasonable choices. We recommend further evaluation via simulation studies in real applications. For illustration purposes, we proceed with $\gamma=0.9$. 

Suppose that we observe 32 events in the new trial, the corresponding $w_{EB}$ is 0.5.
```{r}
wEB(obj2, 32)
```

We can then construct the rMAP prior with the empirical Bayes weight, and derive the posterior distribution. Further Bayesian estimation and inference will be based on the posterior distribution, which is also a weighted mixture of Gamma distributions. 
```{r}
rmap <- robustify(map_hat, weight=0.5, mean=0.38, n=1)
(postmix_rmap <- postmix(rmap, n = adt$TotExp[10] , m = adt$NEvent[10]/adt$TotExp[10]))
```

## Appendix: Full Data
```{r, eval=T}
rdt %>% kbl(caption="Number of Events") %>% 
  kable_classic(full_width = F, html_font = "Cambria", latex_options = "HOLD_position")
Edt %>% kbl(caption="Total Exposure (in Years)") %>% 
  kable_classic(full_width = F, html_font = "Cambria", latex_options = "HOLD_position")
```


## References
[1] Zhang, H., Shen, Y., Li, J., Ye, H., Chiang, AY. (2023+). Adaptively Leveraging External Data with Robust Meta-Analytical-Predictive Prior Using Empirical Bayes. 

[2] Roychoudhury, S., Neuenschwander, B. (2020). Bayesian leveraging of historical control data for a clinical trial with time-to-event en *Statistics in Medicine*, 39(7), 984-995.
